Conjugate Gradient, Boosting, and Herding: \\ Connections and Application

### 2014-02-05 arXiv/boosted-mirror-descent.tex
notation fixes

Let $f$ be a real function whose domain is a normed inner product space. The {\em convex conjugate} $f^{*}$ of $f$ is defined as
\[
f^{*}(x') \eqdef \max_{x} \ip{x}{x'} - f(x).
\]
If $f$ is proper, lower semi-continuous, and convex, then by the 
Fenchel-Moreau theorem $f = f^{**}$, i.e.
\[
f(x) = \max_{x'} \ip{x}{x'} - f^{*}(x').
\]
Throughout the paper, we will assume without further notice that functions are proper and lower semi-continuous when required. The (sub)gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
\[
\partial f^{*}(x') &\in \argmax_{x} \ip{x}{x'} - f(x) \\
\partial f(x) &\in \argmax_{x'} \ip{x}{x'} - f^{*}(x').
\]

@@

We can now define the \bmd  algorithms. Let $X$ and $Y$ be convex subsets 
of the Banach spaces $\mcX$ and $\mcY$, respectively, and let 
$\blf{\cdot}{\cdot}$ be a bilinear map on $\mcX \times \mcY$; typically 
$\blf{\cdot}{\cdot}$ will correspond to an inner product.
% but we will use the bracket notation for convenience even when this is not the case.
Consider the two-argument loss function $L : X \times Y \to \reals$ 
defined as
\[
L(x,y) = \psi(x) + \blf{x}{y} - R(y)
\]
for which we wish to find arguments that obtain $L_{*} = \min_{x}\max_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
\[
L(x)
&= \psi(x) + \max_{y \in \Theta} \left\{\blf{x}{y} - R(y) \right\} \\
&= \psi(x) + R^{*}(x).
\]
We can think of $\psi$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.  The pair of regularization functions $\psi$ and $R$ define a flexible class of functions to optimize, as seen from the following examples:
\begin{enumerate}
\item Let $X$ be the probability simplex over a space $\sS$, let $\phi : \sS \to \reals^d$ be a collection 
      of statistics, and let $Y$ be the unit $L^1$ ball in $\reals^d$. Let $\psi(x) = 0$, 
      $\blf{x}{y} = y^T\bE_{v \sim x}[\phi(v)]$, and 
      $R(y) = \blf{x_0}{y}$. Then $L(x)$ is the 
      \emph{maximum mean discrepancy} between $x$ and $x_0$ relative to $\phi$.
\item \NA{JHH: I don't remember what this example what supposed to be}
	  Let $\psi(x) = 0$ and $R(y) = S^*(i^{-1}(y))$ where $S^*$ is the 
      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(x) = S(x)$ 
      as long as $i^{-1}(\Theta) = \sB_{U}$.
\item Let $X = Y = \reals^d$. Let $\blf{x}{y} = \ip{x}{y}$, $\psi(x) = \|x\|_1$, $R(x) = \blf{x_0}{y} + \frac{1}{2} \|y\|_2^2$. 
      Then $L(x) = \|x\|_1 + \frac{1}{2} \|x-x_0\|_2^2$. 
\end{enumerate}

We will assume 
throughout that $\psi$ and $R$ are both convex functions, and furthermore that 
$\arg\max_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
computed for all values of $y$.

@@

Then 
the {\em primal boosted mirror descent} (\primal) algorithm is:
\begin{enumerate}
\item $x_1 \in \arg\min_x \psi(x)$
\item $y_{t} \in \arg\max_{y \in \Theta} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
\item $x_{t+1} \in \arg\min_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
\end{enumerate}
As long as $\psi$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
bound (see Corollary~\ref{cor:method-1}):
\begin{equation}
\sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
\end{equation}
In other words, $\bar{x}_T$ is close to being a global minimum of $L(x)$.
However, it is often the case that $\psi$ is not strongly convex but $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
\begin{enumerate}
\item $y_1 \in \arg\max_{y} R(y)$
\item $x_t \in \arg\min_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
\item $y_{t+1} \in \arg\max_{y \in \Theta} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
\end{enumerate}
We then obtain the following bound 
when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
\[ \sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]

>>>>

Let $f$ be a real function whose domain is a normed inner product space. The {\em convex conjugate} $f^{*}$ of $f$ is defined as
\[
f^{*}(x') \eqdef \sup_{x} \ip{x}{x'} - f(x).
\]
If $f$ is proper, lower semi-continuous, and convex, then by the 
Fenchel-Moreau theorem $f = f^{**}$, i.e.
\[
f(x) = \sup_{x'} \ip{x}{x'} - f^{*}(x').
\]
Throughout the paper, we will assume without further notice that functions are proper and lower semi-continuous when required. The (sub)gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
\[
\partial f^{*}(x') &\in \arg\sup_{x} \ip{x}{x'} - f(x) \\
\partial f(x) &\in \arg\sup_{x'} \ip{x}{x'} - f^{*}(x').
\]

@@

We can now define the \bmd  algorithms. Let $X$ and $Y$ be convex subsets 
of the Banach spaces $\mcX$ and $\mcY$, respectively, and let 
$\blf{\cdot}{\cdot}$ be a bilinear map on $\mcX \times \mcY$; typically 
$\blf{\cdot}{\cdot}$ will correspond to an inner product.
% but we will use the bracket notation for convenience even when this is not the case.
Consider the two-argument loss function $L : X \times Y \to \reals$ 
defined as
\[
L(x,y) = \psi(x) + \blf{x}{y} - R(y)
\]
for which we wish to find arguments that obtain $L_{*} = \inf_{x}\sup_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
\[
L(x)
&= \psi(x) + \sup_{y \in Y} \left\{\blf{x}{y} - R(y) \right\} \\
&= \psi(x) + R^{*}(x).
\]
We can think of $\psi$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.  The pair of regularization functions $\psi$ and $R$ define a flexible class of functions to optimize, as seen from the following examples:
\begin{enumerate}
\item Let $X$ be the probability simplex over a space $\sS$, let $\phi : \sS \to \reals^d$ be a collection 
      of statistics, and let $Y$ be the unit $L^1$ ball in $\reals^d$. Let $\psi(x) = 0$, 
      $\blf{x}{y} = y^T\bE_{v \sim x}[\phi(v)]$, and 
      $R(y) = \blf{x_0}{y}$. Then $L(x)$ is the 
      \emph{maximum mean discrepancy} between $x$ and $x_0$ relative to $\phi$.
\item \NA{JHH: I don't remember what this example what supposed to be}
	  Let $\psi(x) = 0$ and $R(y) = S^*(i^{-1}(y))$ where $S^*$ is the 
      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(x) = S(x)$ 
      as long as $i^{-1}(Y) = \sB_{U}$.
\item Let $X = Y = \reals^d$. Let $\blf{x}{y} = \ip{x}{y}$, $\psi(x) = \|x\|_1$, $R(x) = \blf{x_0}{y} + \frac{1}{2} \|y\|_2^2$. 
      Then $L(x) = \|x\|_1 + \frac{1}{2} \|x-x_0\|_2^2$. 
\end{enumerate}

We will assume 
throughout that $\psi$ and $R$ are both convex functions, and furthermore that 
$\arg\sup_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
computed for all values of $y$.

@@

Then 
the {\em primal boosted mirror descent} (\primal) algorithm is:
\begin{enumerate}
\item $x_1 \in \arg\inf_x \psi(x)$
\item $y_{t} \in \arg\sup_{y \in Y} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
\item $x_{t+1} \in \arg\inf_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
\end{enumerate}
As long as $\psi$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
bound (see Corollary~\ref{cor:method-1}):
\begin{equation}
\sup_{y \in Y} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
\end{equation}
In other words, $\bar{x}_T$ is close to being a global minimum of $L(x)$.
However, it is often the case that $\psi$ is not strongly convex but $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
\begin{enumerate}
\item $y_1 \in \arg\sup_{y} R(y)$
\item $x_t \in \arg\inf_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
\item $y_{t+1} \in \arg\sup_{y \in Y} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
\end{enumerate}
We then obtain the following bound 
when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
\[ \sup_{y \in Y} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]
